### 2022-04-18 Draft Report/Draft Project.Rmd
Added EJSCreen code and Filtered CSV's (see line 162 and then 200/225/250)

---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman
---

```{r, initialize}
# Set your working directory
getwd()
knitr::opts_knit$set(root.dir = '/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Project/')
getwd()
```

## Potential research questions!!!

-   Write what you're thinking about!
-   <https://www.thetransportpolitic.com/2022/01/24/openings-and-construction-starts-planned-for-2022/>
-   <https://theintercept.com/2022/02/12/wildfires-prisons-climate-california/>
-   <https://cseveren.github.io/files/racialized_difference_commuting_20220320.pdf> <https://github.com/OpenFloodAI/Climate-Change-Datasets> <https://github.com/firstlookmedia/climate-and-punishment-data> \newpage-

```{=tex}
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage
```
```{r setup, include=FALSE}
# Load your packages
library(agricolae)
library(corrplot)
library(cowplot)
library(dataRetrieval)
library(extrafont)
library(extrafontdb)
library(ggpubr)
library(ggthemes)
library(hrbrthemes)
library(Kendall)
library(leaflet)
library(lubridate)
library(mapview)
library(rvest)
library(sf)
library(tidycensus)
library(tidyverse)
library(trend)
library(tseries)
library(viridis)
library(zoo)

# Set your ggplot theme
#Disable on-the-fly projections
sf::sf_use_s2(FALSE)
#Fix Mapview
mapviewOptions(fgb = FALSE)

Theme2 <- theme_ipsum()+
  theme(legend.position = "bottom",
        legend.key = element_rect(fill = "white", colour = "black"),legend.direction = "horizontal",
        legend.title = element_text(face = "bold"))
theme_set(Theme2)
```

# Rationale and Research Questions

# Dataset Information

## Methodology

### Gather Data Sources

```{r load datasources}
# Load your datasets
home_path <- '~/Desktop/Duke MPP/Environ Data /872 Final Project'

NRICali <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_California/NRI_Table_CensusTracts_California.csv',sep=""), stringsAsFactors = TRUE)
NRINorCar <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_NorthCarolina/NRI_Table_CensusTracts_NorthCarolina.csv',sep=""), stringsAsFactors = TRUE)
NRIFlor <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/NRI_Table_CensusTracts_Florida/NRI_Table_CensusTracts_Florida.csv',sep=""), stringsAsFactors = TRUE)
EJScreen20 <- read.csv(paste(home_path,'/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2020_StatePctile.csv',sep=""),stringsAsFactors = TRUE)
#colnames(EJScreen20)
#colnames(EJScreen21)
#EJScreen21 <- read.csv('~/Desktop/Duke MPP/Environ Data /872 Final Project/AmanfuJaniRoberton_ENV872_EDA_FinalProject/Potential Data Sources/EJSCREEN_2021_StatePctile_Tracts.csv',stringsAsFactors = TRUE)
```

####EJSCREEN Gathered From: 2020 #### National Risk Index Gathered from: <https://hazards.fema.gov/nri/data-resources> North Carolina And from <https://gaftp.epa.gov/EJSCREEN/2020/> for the EJ Screen Data ##\# Clean Data Sources

Filter and pipe each to just the variables of interest: location, titles, variables 1 and 2

```{r initial filter}
NRICali2 <- NRICali %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
                !starts_with("LNDS"),!starts_with("VLCN"))

NRIFlor2 <- NRIFlor %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
                !starts_with("LNDS"),!starts_with("VLCN"))

NRINorCar2 <- NRINorCar %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,
         ends_with("RATNG"),ends_with("RISKS"),ends_with("RISKR"),ends_with("EVENTS"),
         ends_with("SCORE"),ends_with("EALS"),ends_with("EALR")) %>%
         select(!starts_with("LTNG"),!starts_with("AVLN"),!starts_with("HAIL"),
                !starts_with("LNDS"),!starts_with("VLCN"))

EJScreen20 <- EJScreen20 %>%
  filter(grepl('Florida|California|North Carolina',STATE_NAME))
```

Questions:(NC, FL, CA) 1. 1960 Housing (size of datapoint) v Coastal Flood risk (Color on map) 2. Percent Low Income v Flood Risk --3? PM2.5 (size of datapoint) v Wildfire Risk (color)--

```{r risk filters}
#flood risks
NRICaliFlood <- NRICali2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
NRIFlorFlood <- NRIFlor2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))
NRINorCarFlood <- NRINorCar2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))

#fire risks
NRICaliFire <- NRICali2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
NRIFlorFire <- NRIFlor2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
NRINorCarFire <- NRINorCar2 %>%
  select(OID_,NRI_ID,STATE,COUNTY,TRACTFIPS,POPULATION,starts_with("WFIR"))
```

```{r EJScreen Filters}
EJ_filter <- EJScreen20 %>%
  select(ID,STATE_NAME,contains("INCPCT"),contains("HSPCT"),LINGISOPCT,PRE1960PCT,contains("LDPNT"),contains("PM25"))

EJ_California <- EJ_filter %>%
    filter(STATE_NAME=="California")

EJ_NorCar <- EJ_filter %>%
  filter(STATE_NAME=='North Carolina')

EJ_Florida <-EJ_filter %>%
  filter(STATE_NAME=="Florida")

## COMMENT THESE OUT THEN READ THEM IN
write.table(EJ_California,'./Potential Data Sources/Census Tracts/EJScreen_California.csv',col.names=TRUE,sep=",")
write.table(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv',col.names=TRUE,sep=",")
write.table(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_Florida.csv',col.names=TRUE,sep=",")
###

# EJ_NorCar <- EJScreen20 %>%
#   filter(STATE_NAME=='North Carolina')
# 
# EJ_Florida <-EJScreen20 %>%
#   filter(STATE_NAME=="Florida")

#write.csv(EJ_NorCar,'./Potential Data Sources/Census Tracts/EJScreen_NC.csv',col.names)
#write.csv(EJ_Florida,'./Potential Data Sources/Census Tracts/EJScreen_FLorida.csv')
```

### Present Descriptive Statistics and Maps

Add in Shape Files <https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html>

```{r census tracts}
california_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_cali_tract_500k/cb_2020_06_tract_500k.shp',)
norcar_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_norcar_tract_500k/cb_2020_37_tract_500k.shp')
florida_tract <- st_read('./Potential Data Sources/Census Tracts/cb_2020_flor_tract_500k/cb_2020_12_tract_500k.shp')
#mapview(california_tract,zcol="AWATER")
```

```{r california file filtering}
california_tract <- california_tract %>%
  select(GEOID,NAMELSAD,NAMELSADCO,ALAND,AWATER,geometry) %>%
  mutate(TRACTFIPS = as.double(substr(california_tract$GEOID,2,15)))
NRICaliFire <- NRICaliFire %>%
  select(TRACTFIPS,POPULATION,starts_with("WFIR"))
NRICaliFlood <- NRICaliFlood %>%
  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))

# See GEOID Structure, <https://www.census.gov/programs-surveys/geography/guidance/geo-identifiers.html> all the other stuff is Census Tracts not BlockGroups
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_California.csv'
names(EJ_California)[names(EJ_California) == 'ID'] <- 'BLOCKID'
EJ_California <- EJ_California %>%
  mutate(TRACTID = as.double(substr(EJ_California$BLOCKID,1,10)))

#now we convert into averages for the whole census tract. Imprecise but we've gotta do it
EJ_Cali <- EJ_California %>%
  select(!starts_with("T_")) %>%
  group_by(TRACTID) %>%
  summarize(across(!where(is.factor),mean))
```

```{r florida file filtering}
florida_tract <- florida_tract %>%
  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
  mutate(TRACTFIPS = as.double(substr(florida_tract$GEOID, 1,15)))
NRIFlorFire <-NRIFlorFire %>%
  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
NRIFlorFlood <- NRIFlorFlood %>%
  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))

# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_Florida.csv'
EJScreen_Florida <- EJ_Florida
names(EJScreen_Florida)[names(EJScreen_Florida) == 'ID'] <- 'BLOCKID' 
EJScreen_Florida <- EJScreen_Florida %>% mutate(TRACTID = as.double(substr(EJScreen_Florida$BLOCKID,1,11)))

EJ_Flor <- EJScreen_Florida %>%
select(!starts_with("T_")) %>%
group_by(TRACTID) %>%
summarize(across(!where(is.factor), mean))
```

```{r north carolina file filtering}
norcar_tract <- norcar_tract %>%
  select(GEOID,NAMELSAD, NAMELSADCO, ALAND, AWATER, geometry) %>%
  mutate(TRACTFIPS = as.double(substr(norcar_tract$GEOID, 1,15)))
NRINorCarFire <-NRINorCarFire %>%
  select(TRACTFIPS,POPULATION,starts_with ("WFIR"))
NRINorCarFlood <- NRINorCarFlood %>%
  select(TRACTFIPS,POPULATION,starts_with("CFLD"),starts_with("RFLD"))

# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
# READ IN THE CSV FILE FOR './Potential Data Sources/Census Tracts/EJScreen_NorthCarolina.csv'
EJScreen_norcar <- EJ_NorCar
names(EJScreen_norcar)[names(EJScreen_norcar) == 'ID'] <- 'BLOCKID' 
EJScreen_norcar <- EJScreen_norcar %>% mutate(TRACTID = as.double(substr(EJScreen_norcar$BLOCKID,1,11)))

EJ_NorCar <- EJScreen_norcar %>%
select(!starts_with("T_")) %>%
group_by(TRACTID) %>%
summarize(across(!where(is.factor), mean))
```

<https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html>

```{r longitude latitute for counties}
#https://www.census.gov/geographies/reference-files/time-series/geo/gazetteer-files.2020.html
gazetteer <- read.delim('./Potential Data Sources/Census Tracts/2021_Gaz_tracts_national.txt',stringsAsFactors = TRUE)

ca_gazetteer <- gazetteer %>%
  filter(USPS=="CA")

ca_gazetteer <- ca_gazetteer %>%  
  select(!c(ALAND_SQMI,AWATER_SQMI))%>%
  mutate(TRACTID = ca_gazetteer$GEOID)

fl_gazetteer <- gazetteer %>%
  filter(USPS=="FL")

fl_gazetteer <- fl_gazetteer %>%
  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
  mutate(TRACTID = fl_gazetteer$GEOID)

nc_gazetteer <- gazetteer %>%
  filter(USPS=="NC")

nc_gazetteer <- nc_gazetteer %>%
  select(!c(ALAND_SQMI, AWATER_SQMI)) %>%
  mutate(TRACTID = nc_gazetteer$GEOID)
```

```{r attempts to merge}
#California
CA_NRI_fire_shp <- left_join(california_tract,NRICaliFire,by="TRACTFIPS")
CA_NRI_flood_shp <- left_join(california_tract,NRICaliFlood,by="TRACTFIPS")

CA_EJ_df <- left_join(ca_gazetteer,EJ_Cali,by="TRACTID")
CA_EJ_shp <- CA_EJ_df %>%
  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)

#Florida
FL_NRI_fire_shp <- left_join(florida_tract,NRIFlorFire,by="TRACTFIPS")
FL_NRI_flood_shp <- left_join(florida_tract,NRIFlorFlood,by="TRACTFIPS")

FL_EJ_df <- left_join(fl_gazetteer,EJ_Flor,by="TRACTID")
FL_EJ_shp <- FL_EJ_df %>%
  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)

#North Carolina
NC_NRI_fire_shp <- left_join(norcar_tract,NRINorCarFire,by="TRACTFIPS")
NC_NRI_flood_shp <- left_join(norcar_tract,NRINorCarFlood,by="TRACTFIPS")

NC_EJ_df <- left_join(nc_gazetteer,EJ_NorCar,by="TRACTID")
NC_EJ_shp <- NC_EJ_df %>%
  st_as_sf(coords = c('INTPTLONG','INTPTLAT'), crs=4269)
```

```{r simple graphs CA}
ggplot(CA_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")

ggplot(CA_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
ggplot(CA_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
```

```{r simple graphs FL}
ggplot(FL_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")

ggplot(FL_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
ggplot(FL_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
```

```{r simple graphs NC}
ggplot(NC_EJ_df,aes(y=LOWINCPCT,x=PM25))+geom_point(aes(color=LINGISOPCT))+geom_smooth(method="lm")+labs(title="PM25 v Low Income %",y="% Low Income Census Tract",x="PM25 Concentration")

ggplot(NC_NRI_fire_shp,aes(y=POPULATION,x=WFIR_RISKR))+geom_boxplot()+coord_flip()+labs(title="Fire Risk versus Population",y="Fire Risk")
ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=CFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Coastal Flood Risk versus Population",y="Coastal Flood Risk")
ggplot(NC_NRI_flood_shp,aes(y=POPULATION,x=RFLD_RISKR))+geom_boxplot()+coord_flip()+labs(title="Riverine Flood Risk versus Population",y="Riverine Flood Risk")
```

```{r simple mapview tests}
#CA
mapview(CA_NRI_fire_shp,zcol="POPULATION")+ mapview(CA_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(CA_EJ_shp,zcol="LOWINCPCT")
#comment
#FL
mapview(FL_NRI_fire_shp,zcol="POPULATION")+ mapview(FL_NRI_flood_shp,zcol="RFLD_RISKR") +mapview(FL_EJ_shp,zcol="LOWINCPCT")

#NC
mapview(NC_NRI_fire_shp,zcol="POPULATION")+mapview(NC_NRI_flood_shp,zcol="RFLD_RISKR")+mapview(NC_EJ_shp,zcol="LOWINCPCT")
```

```{r mapview tests CA}
mapview(list(CA_NRI_fire_shp,CA_EJ_shp),
        zcol=c("WFIR_RISKS","PRE1960PCT"),
        layer.name=list("Fire Risk","Relative Building Ages"))

mapview(list(CA_NRI_flood_shp,CA_EJ_shp),
        zcol=list("CFLD_RISKS","PRE1960PCT"),
        layer.name=list("Coastal Flood Risk","Relative Building Ages"),
        col.regions=list(mapviewPalette("mapviewVectorColors"),mapviewPalette("mapviewTopoColors")),
        alpha.regions=.5)

mapview(CA_NRI_fire_shp,zcol="WFIR_RISKS")
mapview(CA_EJ_shp,zcol="LOWINCPCT",alpha.regions=.5,col.regions= mapviewPalette("mapviewTopoColors"))
mapview(CA_EJ_shp,zcol="PRE1960PCT")

mapview(list(NC_NRI_fire_shp,NC_EJ_shp),
        zcol=c("WFIR_RISKS","PRE1960PCT"),
        layer.name=list("Fire Risk","Relative Building Ages"))

# mapshot(m, file = paste0(getwd(), "/map.png"),
#           remove_controls = c("homeButton", "layersControl"))
```

### Time Series

\newpage

# Exploratory Analysis

\newpage

# Analysis

## Question 1: \<insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2:

\newpage

# Summary and Conclusions

\newpage

# References

\<add references here if relevant, otherwise delete this section>